---
# title: "CryptoIndicator Report"
# author: "by Filippo Franchini"

output: 
  pdf_document:
    number_sections: no
    keep_tex: true
    extra_dependencies: ["float"]

header-includes:
  - \usepackage{makecell}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \lhead{}
  - \fancyhead[R]{}
  - \fancyhead[C]{Crypto Report generated by CryptoIndicator R-Package -- \today}
  - \fancyfoot[L]{\begin{minipage}[c]{6cm}Created by Filippo Franchini \\ Senior Technical Advisor \\ filippoweb3 on X/LinkedIn\end{minipage}}
  - \fancyfoot[R]{\begin{minipage}[c]{2cm}\end{minipage}}
  - \pagenumbering{gobble}
  - \pagestyle{fancy}
  - \setlength{\footskip}{35pt}
  - \usepackage{xcolor}
  - \definecolor{linkblue}{RGB}{0, 102, 204}
  - \newcommand{\ulink}[2]{\textcolor{linkblue}{\underline{\href{#1}{#2}}}}
  - \usepackage{multicol}

vignette: >
  %\VignetteIndexEntry{report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
font_size: 11pt
geometry: left=1cm,right=1cm,top=2cm,bottom=3cm
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.pos = "h", 
  out.extra = "",
  dpi = 300
)
```

```{r setup, include = FALSE}
library(knitr)
library(dplyr)
library(CryptoIndicator)
library(kableExtra)
library(ggplot2)
library(patchwork)
library(zoo)
library(quantmod)

devtools::load_all()

start_date <- "2025-01-01"

# On-chain

filter_by_date <- function(data, date_col, start, end) {
    if (is.null(data) || nrow(data) == 0) return(data)
    data[data[[date_col]] >= start & data[[date_col]] <= end, ]
  }

onchain <- result$indicators$onchain
df <- filter_by_date(onchain, "timestamp", start_date, Sys.Date())

# Volatility

vol <- result$indicators$volatility

regime_data <- data.frame(
  date = result$indicators$onchain$timestamp,
  regime = vol$regime
  )

# Create a colored bar plot for regimes
regime_data$color <- case_when(
  regime_data$regime == "Extreme High" ~ "red",
  regime_data$regime == "High" ~ "orange",
  regime_data$regime == "Low" ~ "lightgreen",
  regime_data$regime == "Extreme Low" ~ "green",
  TRUE ~ "gray"
  )

regime_data <- filter_by_date(regime_data, "date", start_date, Sys.Date())

df1 <- data.frame(
      date = onchain$timestamp,
      short_vol = vol$volatility$short * 100,
      medium_vol = vol$volatility$medium * 100
    )
df1 <- filter_by_date(df1, "date", start_date, Sys.Date())

df2 <- data.frame(
      date = onchain$timestamp,
      breakout = vol$signals$vol_breakout,
      compression = vol$signals$vol_compression
    )

    # Create signals plot
    df2$signal <- ifelse(df2$breakout, "Breakout",
                        ifelse(df2$compression, "Compression", "Normal"))
    df2$color <- ifelse(df2$breakout, "red",
                       ifelse(df2$compression, "green", "gray"))
    
df2 <- filter_by_date(df2, "date", start_date, Sys.Date())

# Derivatives

deriv <- result$raw_data$derivatives
df3 <- filter_by_date(deriv$data, "date", start_date, Sys.Date())
df3$positive <- df3$avg_funding_rate > 0

df4 <- data.frame(
      date = deriv$data$date,
      oi_percentile = deriv$oi_percentile * 100,
      funding_percentile = deriv$funding_percentile * 100
    )

#df4 <- filter_by_date(df4, "date", "2025-01-01", Sys.Date())

macro <- result$raw_data$macro
m2_data <- data.frame(
  date = macro$liquidity$m2_us$date,
  m2_yoy = macro$liquidity$m2_yoy * 100)
m2_data <- filter_by_date(m2_data, "date", start_date, Sys.Date())
m2_data <- m2_data[!is.na(m2_data$m2_yoy), ]

dxy <- macro$dollar$dxy
dxy_df <- data.frame(
  date = as.Date(index(dxy)),
  dxy = as.numeric(Cl(dxy)))

dxy_df <- filter_by_date(dxy_df, "date", start_date, Sys.Date())
```

\pagenumbering{arabic}

\huge \textbf{CryptoIndicator Report}  
\vspace{0.1cm}  
\large \textbf{Explore more with \ulink{https://filippoweb3.shinyapps.io/crypto-indicator-demo/}{CryptoIndicator App}}  
\vspace{0.5cm}  

\normalsize

Report generated by \ulink{https://github.com/filippoweb3/crypto-indicator}{CryptoIndicator R-Package} on `r Sys.Date()`. For detailed methodologies and framework mechanics, see the package README and function documentation.  

**Main sources:** CoinMarketCap, Binance, FRED, Whale Alert, Google Trends  
**Asset:** Bitcoin  
**Risk Level:** \textcolor{orange}{YELLOW} (Elevated Risk)  
**Market Regime:** Bear Market (Accumulation Phase)

\begin{multicols}{2}

\subsection*{Market Overview}

As of February 2026, Bitcoin presents a \textbf{confluence of contradictory signals} that characterizes a market transition zone. The framework's risk-off indicator is YELLOW, reflecting elevated risk factors including a negative macro composite and crowded long positioning in derivatives markets. However, beneath this surface, multiple on-chain metrics are flashing historically reliable accumulation signals.

The \textbf{MVRV ratio} at `r round(result$signals$onchain$mvrv$value, 2)` places Bitcoin firmly in bear market territory, trading below its realized price; a condition that has historically preceded major accumulation phases. This is reinforced by the \textbf{Puell Multiple} at `r round(result$signals$onchain$puell$value, 2)`, indicating miner capitulation (selling at a loss), a phenomenon that has consistently marked cycle bottoms in 2015, 2018, and 2022. The \textbf{NVT ratio} at `r round(result$signals$onchain$nvt$value, 2)` suggests the network is undervalued relative to its transaction volume, further supporting the fundamental value thesis.

\subsection*{Derivatives Market Dynamics}

Despite the constructive on-chain picture, derivatives markets tell a cautionary tale. Funding rates remain neutral, but the framework detects \textbf{crowded long positioning} (OI percentile `r round(result$signals$derivatives$oi_percentile * 100, 1)`%, funding percentile `r round(result$signals$derivatives$funding_percentile * 100, 1)`%). This combination—crowded longs without extreme funding, suggests a market that is positioned bullishly but not yet at speculative excess. The warning is clear: any downside move could trigger a liquidation cascade, making the market vulnerable to sharp corrections.

The crowded long signal serves as a near-term headwind that may delay or mute the upside potential suggested by on-chain metrics. This is typical of market bottoms where "smart money" accumulates while latecomers remain over-leveraged.

\subsection*{Macro Environment}

The macro backdrop remains challenging. With \textbf{composite risk score at `r round(result$signals$macro$composite_risk, 2)`} and a bearish bias, traditional financial conditions are not yet supportive of a sustained crypto rally. High interest rates (risk-off regime) and neutral dollar strength create headwinds, while M2 liquidity remains stable, neither expanding nor contracting significantly.

This macro environment suggests that any near-term upside may be capped, and the market may need to consolidate or drift lower before a sustained recovery can begin. The historical pattern shows that Bitcoin often bottoms while macro conditions remain poor, then rallies as liquidity conditions improve.

\subsection*{Volatility and Positioning}

Volatility regimes currently favor \textbf{trend following} (low volatility environment), yet the optimal strategy is identified as \textbf{mean reversion}, a seeming contradiction that reflects the market's transitional nature. Low volatility with compressed ranges typically precedes significant moves, and the crowded long positioning suggests the next major move could be triggered by a liquidation event.

Whale activity provides a bright spot: \textbf{`r result$signals$positioning$whale_flow`} indicate that large holders are moving coins to cold storage, a classic accumulation signal. This aligns with the on-chain metrics showing value territory. Meanwhile, retail sentiment via Google Trends sits at an extremely low `r round(result$signals$positioning$sentiment$score, 1)`, historically a contrarian bullish signal indicating that \textit{no one is left to sell}.

\subsection*{Strategic Implications}

The framework's medium-term trend remains \textbf{`r result$decisions$medium_term$trend$direction`} (score `r result$decisions$medium_term$trend$score`/4) with a recommended allocation of `r result$decisions$medium_term$position_sizing$recommended_allocation_pct`%, while the long-term strategic view is to \textbf{`r result$decisions$long_term$strategic$core_position`}. This multi-timeframe perspective is crucial: near-term caution due to crowded positioning, medium-term optimism based on valuation, and long-term conviction from cycle analysis.

The current market regime is classified as \textbf{`r result$decisions$regime_allocation$current_regime`} (accumulation phase), with a tactical approach of mean reversion—buying dips and selling rallies within a range. Hedges are recommended to manage the elevated risk profile. The most prudent strategy is to scale into positions gradually, maintaining powder for potential further downside while building a core position at historically attractive valuations.

\end{multicols}

\subsection*{Technical Charts}

```{r, echo=F, fig.height=9, fig.width=8, fig.align="center", warning=FALSE, message=FALSE}
p1 <- ggplot(df, aes(x = timestamp, y = mvrv)) +
      geom_line(color = "darkgray", linewidth = 0.5) +
      geom_hline(yintercept = c(0.7, 0.9, 1.75),
                 linetype = "dashed", color = "red", alpha = 0.5) +
      annotate("rect", xmin = min(df$timestamp), xmax = max(df$timestamp),
               ymin = 0.7, ymax = 0.9, alpha = 0.1, fill = "orange") +
      annotate("text", x = min(df$timestamp), y = 0.65, label = "Extreme Bottom",
               hjust = 0, size = 3, color = "darkgreen") +
      annotate("text", x = min(df$timestamp), y = 1.8, label = "Extreme Top",
               hjust = 0, size = 3, color = "red") +
      labs(title = "MVRV Ratio",
           x = "Date", y = "MVRV") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p2 <- ggplot(df, aes(x = timestamp, y = puell_multiple)) +
      geom_line(color = "darkgray", linewidth = 0.5) +
      geom_hline(yintercept = c(0.7, 2),
                 linetype = "dashed", color = "red", alpha = 0.5) +
      annotate("text", x = min(df$timestamp), y = 0.6, label = "Miner Capitulation",
               hjust = 0, size = 3, color = "darkgreen") +
      annotate("text", x = min(df$timestamp), y = 2.1, label = "Miner Selling",
               hjust = 0, size = 3, color = "red") +
      labs(title = "Puell Multiple",
           x = "Date", y = "Puell Multiple") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p3 <- ggplot(df, aes(x = timestamp, y = nvts)) +
      geom_line(color = "darkgray", linewidth = 0.5) +
      geom_hline(yintercept = c(30, 70),
                 linetype = "dashed", color = "red", alpha = 0.5) +
      annotate("text", x = min(df$timestamp), y = 25, label = "Undervalued",
               hjust = 0, size = 3, color = "darkgreen") +
      annotate("text", x = min(df$timestamp), y = 75, label = "Overvalued",
               hjust = 0, size = 3, color = "red") +
      labs(title = "NVT Ratio",
           x = "Date", y = "NVT") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p4 <- ggplot(regime_data, aes(x = date, y = 1, fill = regime)) +
      geom_tile(height = 0.8) +
      scale_fill_manual(values = c("Extreme High" = "red",
                                   "High" = "orange",
                                   "Low" = "lightgreen",
                                   "Extreme Low" = "green")) +
      labs(title = "Volatility Regimes",
           x = "Date", y = "", fill = "Regime") +
      theme_minimal() +
      theme(axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            plot.title = element_text(hjust = 0.5),
            legend.position = "right",
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p5 <- ggplot(df1, aes(x = date)) +
      geom_line(aes(y = short_vol, color = "Short-term (7d)"), linewidth = 0.5) +
      geom_line(aes(y = medium_vol, color = "Medium-term (30d)"), linewidth = 0.5, linetype = "dashed") +
      scale_color_manual(values = c("Short-term (7d)" = "darkgray",
                                    "Medium-term (30d)" = "red")) +
      labs(title = "Volatility Comparison",
           x = "Date", y = "Annualized Volatility (%)",
           color = "Horizon") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            legend.position = "right",
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p6 <- ggplot(df2, aes(x = date, y = 1, fill = signal)) +
      geom_tile(height = 0.8) +
      scale_fill_manual(values = c("Breakout" = "red",
                                   "Compression" = "green",
                                   "Normal" = "gray")) +
      labs(title = "Volatility Signals",
           x = "Date", y = "", fill = "Signal Type") +
      theme_minimal() +
      theme(axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            plot.title = element_text(hjust = 0.5),
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p7 <- ggplot(df3, aes(x = date, y = avg_funding_rate * 100)) +
      geom_col(aes(fill = positive), alpha = 0.7) +
      geom_hline(yintercept = 0, linetype = "solid") +
      geom_hline(yintercept = c(0.01, 0.05, -0.01),
                 linetype = "dashed", color = "gray50", alpha = 0.5) +
      scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red")) +
      labs(title = "Funding Rates",
           x = "Date", y = "Funding Rate (%)") +
      theme_minimal() +
      theme(legend.position = "none",
            plot.title = element_text(hjust = 0.5),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9))

p8 <- ggplot(df4, aes(x = date)) +
      geom_line(aes(y = oi_percentile, color = "OI Percentile"), linewidth = 0.5) +
      geom_line(aes(y = funding_percentile, color = "Funding Percentile"),
                linewidth = 0.5, linetype = "dashed") +
      geom_hline(yintercept = c(10, 90), linetype = "dotted", color = "gray50") +
      scale_color_manual(values = c("OI Percentile" = "blue",
                                    "Funding Percentile" = "red")) +
      labs(title = "Percentile Ranks",
           x = "Date", y = "Percentile (%)",
           color = "Metric") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            legend.position = "right",
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9))

p9 <- ggplot(m2_data, aes(x = date, y = m2_yoy)) +
      geom_line(color = "darkgreen", linewidth = 0.5) +
      geom_hline(yintercept = c(0, 5), linetype = "dashed", color = "red") +
      annotate("rect", xmin = min(m2_data$date), xmax = max(m2_data$date),
               ymin = 5, ymax = max(m2_data$m2_yoy, na.rm = TRUE),
               alpha = 0.1, fill = "green") +
      annotate("rect", xmin = min(m2_data$date), xmax = max(m2_data$date),
               ymin = min(m2_data$m2_yoy, na.rm = TRUE), ymax = 0,
               alpha = 0.1, fill = "red") +
      labs(title = "M2 Money Supply Growth",
           x = "Date", y = "YoY % Change") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p10 <- ggplot(dxy_df, aes(x = date, y = dxy)) +
      geom_line(color = "blue", linewidth = 0.5) +
      geom_hline(yintercept = c(90, 100), linetype = "dashed", color = "red") +
      annotate("rect", xmin = min(dxy_df$date), xmax = max(dxy_df$date),
               ymin = 90, ymax = 100, alpha = 0.1, fill = "orange") +
      labs(title = "US Dollar Index (DXY)",
           x = "Date", y = "DXY") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

# First create the left and right stacks
left_stack <- p1 / p2 / p3 / p9 / p10
right_stack <- p4 / p5 / p6 / p7 / p8

# Then combine them side by side and apply layout
combined_plot <- (left_stack | right_stack)

# Display the plot
combined_plot

```
