---
# title: "CryptoIndicator Report"
# author: "by Filippo Franchini"

output: 
  pdf_document:
    number_sections: no
    keep_tex: true
    extra_dependencies: ["float"]

header-includes:
  - \usepackage{makecell}
  - \usepackage{fancyhdr}
  - \pagestyle{fancy}
  - \lhead{}
  - \fancyhead[R]{}
  - \fancyhead[C]{Crypto Report generated by CryptoIndicator R-Package -- \today}
  - \fancyfoot[L]{\begin{minipage}[c]{6cm}Created by Filippo Franchini \\ filippoweb3 on X/LinkedIn\end{minipage}}
  - \fancyfoot[R]{\begin{minipage}[c]{2cm}\end{minipage}}
  - \pagenumbering{gobble}
  - \pagestyle{fancy}
  - \setlength{\footskip}{35pt}
  - \usepackage{xcolor}
  - \definecolor{linkblue}{RGB}{0, 102, 204}
  - \newcommand{\ulink}[2]{\textcolor{linkblue}{\underline{\href{#1}{#2}}}}
  - \usepackage{multicol}

vignette: >
  %\VignetteIndexEntry{report}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
font_size: 7pt
geometry: left=1cm,right=1cm,top=2cm,bottom=3cm
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.pos = "h", 
  out.extra = "",
  dpi = 300
)
```

```{r setup, include = FALSE}
library(knitr)
library(dplyr)
library(CryptoIndicator)
library(kableExtra)
library(ggplot2)
library(patchwork)
library(zoo)
library(quantmod)

devtools::load_all()

start_date <- "2025-01-01"

# On-chain

filter_by_date <- function(data, date_col, start, end) {
    if (is.null(data) || nrow(data) == 0) return(data)
    data[data[[date_col]] >= start & data[[date_col]] <= end, ]
  }

onchain <- result$indicators$onchain
df <- filter_by_date(onchain, "timestamp", start_date, Sys.Date())

# Volatility

vol <- result$indicators$volatility

regime_data <- data.frame(
  date = result$indicators$onchain$timestamp,
  regime = vol$regime
  )

# Create a colored bar plot for regimes
regime_data$color <- case_when(
  regime_data$regime == "Extreme High" ~ "red",
  regime_data$regime == "High" ~ "orange",
  regime_data$regime == "Low" ~ "lightgreen",
  regime_data$regime == "Extreme Low" ~ "green",
  TRUE ~ "gray"
  )

regime_data <- filter_by_date(regime_data, "date", start_date, Sys.Date())

df1 <- data.frame(
      date = onchain$timestamp,
      short_vol = vol$volatility$short * 100,
      medium_vol = vol$volatility$medium * 100
    )
df1 <- filter_by_date(df1, "date", start_date, Sys.Date())

df2 <- data.frame(
      date = onchain$timestamp,
      breakout = vol$signals$vol_breakout,
      compression = vol$signals$vol_compression
    )

    # Create signals plot
    df2$signal <- ifelse(df2$breakout, "Breakout",
                        ifelse(df2$compression, "Compression", "Normal"))
    df2$color <- ifelse(df2$breakout, "red",
                       ifelse(df2$compression, "green", "gray"))
    
df2 <- filter_by_date(df2, "date", start_date, Sys.Date())

# Derivatives

deriv <- result$raw_data$derivatives
df3 <- filter_by_date(deriv$data, "date", start_date, Sys.Date())
df3$positive <- df3$avg_funding_rate > 0

df4 <- data.frame(
      date = deriv$data$date,
      oi_percentile = deriv$oi_percentile * 100,
      funding_percentile = deriv$funding_percentile * 100
    )

#df4 <- filter_by_date(df4, "date", "2025-01-01", Sys.Date())

macro <- result$raw_data$macro
m2_data <- data.frame(
  date = macro$liquidity$m2_us$date,
  m2_yoy = macro$liquidity$m2_yoy * 100)
m2_data <- filter_by_date(m2_data, "date", start_date, Sys.Date())
m2_data <- m2_data[!is.na(m2_data$m2_yoy), ]

# FIXED: DXY data extraction for FRED
dxy <- macro$dollar$dxy

if (!is.null(dxy) && nrow(dxy) > 0) {
  dxy_df <- data.frame(
    date = as.Date(dxy$date),  # FRED date column
    dxy = dxy$value             # FRED value column
  )
} else {
  # Fallback
  dxy_df <- data.frame(
    date = seq(as.Date(start_date), Sys.Date(), by = "day"),
    dxy = NA_real_
  )
}

dxy_df <- filter_by_date(dxy_df, "date", start_date, Sys.Date())
```

\pagenumbering{arabic}

\huge \textbf{CryptoIndicator Report}
\vspace{0.1cm}
\large \textbf{Explore more with \ulink{https://filippoweb3.shinyapps.io/crypto-indicator-demo/}{CryptoIndicator App}}
\vspace{0.1cm}

\normalsize

Report generated by \ulink{https://github.com/filippoweb3/crypto-indicator}{CryptoIndicator R-Package} on r Sys.Date(). For detailed methodologies and framework mechanics, see the package README and function documentation.

Main sources: CoinMarketCap, Binance, FRED, Morpho Vaults, Google Trends  
Asset: Bitcoin  
Risk Level: `r result$summary$risk$color` - `r result$summary$risk$level` (`r round(result$summary$risk$composite_score, 2)` composite)  
Market Regime: `r result$summary$regime$name` (`r result$summary$regime$confidence` confidence)

\begin{multicols}{2}

\subsection*{Market Overview}

As of March 2026, Bitcoin presents a \textbf{classic capitulation setup} with the framework's risk-off indicator flashing RED — its highest alert level. This signals extreme market stress requiring immediate defensive positioning. However, beneath this surface, multiple on-chain metrics are converging at historically reliable accumulation zones, creating a tension between short-term safety and long-term opportunity.

The \textbf{MVRV ratio} at `r round(result$signals$onchain$mvrv$value, 2)` places Bitcoin in bear market territory, trading below its realized price. Critically, this is within `r gsub("[^0-9.]", "", result$signals$onchain$mvrv$proximity_warnings[1])`\% of the 0.7 extreme bottom threshold — a level that has marked every major cycle bottom in Bitcoin's history. This proximity warning is the framework's way of signaling that we are in a high-conviction accumulation zone.

The \textbf{Puell Multiple} at `r round(result$signals$onchain$puell$value, 2)` confirms this picture, indicating active miner capitulation (selling at a loss). This phenomenon has consistently marked cycle bottoms in 2015, 2018, and 2022. Miners, as the market's natural sellers, capitulating at current prices suggests we are near peak selling pressure exhaustion. The \textbf{NVT ratio} at `r round(result$signals$onchain$nvt$value, 2)` supports the undervaluation thesis, though it remains in neutral territory.

\subsection*{Risk-Off Signal Analysis}

The RED risk level is driven by a convergence of factors: elevated volatility, a negative macro composite, and critically, extreme sentiment readings. The Fear \& Greed Index at `r result$signals$sentiment$value` registers as Extreme Fear, with the framework correctly interpreting this as a bullish contrarian signal (`r result$signals$sentiment$contrarian_signal`). When retail capitulation reaches this level, it historically precedes trend reversals.

The framework's risk override is actively downgrading all time horizon signals. Short-term trading is disabled ("RISK-OFF"), medium-term allocation is reduced to just 15\%, and long-term strategy shifts to "defensive accumulation" — building positions only on severe dips. This hierarchical approach ensures safety takes precedence over opportunity.

\subsection*{Derivatives Market Dynamics}

Derivatives markets tell a cautionary tale with limited data confidence (only 28 days of history). Funding rates are bearish, and the OI percentile at `r round(result$signals$derivatives$oi_percentile * 100, 1)`\% indicates elevated open interest despite the price decline. This suggests leveraged longs may still be trapped, creating vulnerability to further liquidation cascades.

The framework's interpretation — "cautious short positioning" — reflects a market where shorts are in control but not yet at extreme levels that would signal a squeeze setup. The low funding percentile (`r round(result$signals$derivatives$funding_percentile * 100, 1)`\%) confirms that funding rates are not yet at capitulation levels seen at previous bottoms.

\subsection*{Macro Environment}

The macro backdrop remains challenging with a composite risk score of `r round(result$signals$macro$composite_risk, 2)` and bearish bias. High interest rates (risk-off regime) continue to create headwinds for all risk assets, while M2 liquidity remains stable — neither contracting sharply (which would worsen conditions) nor expanding (which would fuel a rally). The dollar index remains in neutral territory, offering no clear directional cue.

This macro environment suggests that any sustained recovery may require a shift in broader financial conditions. Historically, Bitcoin bottoms while macro remains poor, then rallies as liquidity improves — suggesting we may be in the "bottoming" rather than "rallying" phase.

\subsection*{Volatility and Whale Positioning}

Volatility regimes are currently classified as "High" with an "Active Trading" environment, yet the optimal strategy shows "No Clear Volatility Signal" — a reflection of the market's transitional nature. The volatility comparison shows short-term volatility (77.8th percentile) significantly above medium-term (47.3rd percentile), indicating we are in a volatile period that may be exhausting itself.

Whale activity via Morpho vaults provides crucial intelligence: moderate outflows (`r round(result$signals$positioning$vault_flows$net_flow_pct, 1)`\% of TVL) indicate that large holders are gradually reducing DeFi exposure, a cautious signal. However, whale concentration remains low (HHI of `r round(result$signals$positioning$whale_concentration$hhi, 1)`, competitive), suggesting no systemic risk from concentration. The market is broadly held rather than dominated by a few players.

\subsection*{Strategic Implications}

The framework's multi-timeframe perspective is now crucial: short-term is RISK-OFF (no new positions), medium-term is RISK-OFF with just 15\% allocation, and long-term is defensive accumulation — building a core position only on severe dips. This hierarchy respects the immediate danger while acknowledging the generational value opportunity.

The current bear regime (high confidence) calls for a mean reversion tactical approach — buying dips and selling rallies within a range — with hedges recommended. The recommended strategic allocation of 25-50\% reflects this defensive posture, preserving capital while maintaining exposure to a potential reversal.

The most prudent strategy for March 2026 is clear: preserve capital first, accumulate patiently second. Keep the majority in stablecoins, use limit orders at levels below current prices, and wait for the risk-off signal to moderate before deploying significant capital. The on-chain setup is among the most attractive in years, but the market structure demands patience.

\end{multicols}

\subsection*{Technical Charts}

```{r, echo=F, fig.height=9, fig.width=8, fig.align="center", warning=FALSE, message=FALSE}
p1 <- ggplot(df, aes(x = timestamp, y = mvrv)) +
      geom_line(color = "darkgray", linewidth = 0.5) +
      geom_hline(yintercept = c(0.7, 0.9, 1.75),
                 linetype = "dashed", color = "red", alpha = 0.5) +
      annotate("rect", xmin = min(df$timestamp), xmax = max(df$timestamp),
               ymin = 0.7, ymax = 0.9, alpha = 0.1, fill = "orange") +
      annotate("text", x = min(df$timestamp), y = 0.65, label = "Extreme Bottom",
               hjust = 0, size = 3, color = "darkgreen") +
      annotate("text", x = min(df$timestamp), y = 1.8, label = "Extreme Top",
               hjust = 0, size = 3, color = "red") +
      labs(title = "MVRV Ratio",
           x = "Date", y = "MVRV") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p2 <- ggplot(df, aes(x = timestamp, y = puell_multiple)) +
      geom_line(color = "darkgray", linewidth = 0.5) +
      geom_hline(yintercept = c(0.7, 2),
                 linetype = "dashed", color = "red", alpha = 0.5) +
      annotate("text", x = min(df$timestamp), y = 0.6, label = "Miner Capitulation",
               hjust = 0, size = 3, color = "darkgreen") +
      annotate("text", x = min(df$timestamp), y = 2.1, label = "Miner Selling",
               hjust = 0, size = 3, color = "red") +
      labs(title = "Puell Multiple",
           x = "Date", y = "Puell Multiple") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p3 <- ggplot(df, aes(x = timestamp, y = nvts)) +
      geom_line(color = "darkgray", linewidth = 0.5) +
      geom_hline(yintercept = c(30, 70),
                 linetype = "dashed", color = "red", alpha = 0.5) +
      annotate("text", x = min(df$timestamp), y = 25, label = "Undervalued",
               hjust = 0, size = 3, color = "darkgreen") +
      annotate("text", x = min(df$timestamp), y = 75, label = "Overvalued",
               hjust = 0, size = 3, color = "red") +
      labs(title = "NVT Ratio",
           x = "Date", y = "NVT") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p4 <- ggplot(regime_data, aes(x = date, y = 1, fill = regime)) +
      geom_tile(height = 0.8) +
      scale_fill_manual(values = c("Extreme High" = "red",
                                   "High" = "orange",
                                   "Low" = "lightgreen",
                                   "Extreme Low" = "green")) +
      labs(title = "Volatility Regimes",
           x = "Date", y = "", fill = "Regime") +
      theme_minimal() +
      theme(axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            plot.title = element_text(hjust = 0.5),
            legend.position = "right",
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p5 <- ggplot(df1, aes(x = date)) +
      geom_line(aes(y = short_vol, color = "Short-term (7d)"), linewidth = 0.5) +
      geom_line(aes(y = medium_vol, color = "Medium-term (30d)"), linewidth = 0.5, linetype = "dashed") +
      scale_color_manual(values = c("Short-term (7d)" = "darkgray",
                                    "Medium-term (30d)" = "red")) +
      labs(title = "Volatility Comparison",
           x = "Date", y = "Annualized Volatility (%)",
           color = "Horizon") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            legend.position = "right",
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p6 <- ggplot(df2, aes(x = date, y = 1, fill = signal)) +
      geom_tile(height = 0.8) +
      scale_fill_manual(values = c("Breakout" = "red",
                                   "Compression" = "green",
                                   "Normal" = "gray")) +
      labs(title = "Volatility Signals",
           x = "Date", y = "", fill = "Signal Type") +
      theme_minimal() +
      theme(axis.text.y = element_blank(),
            axis.ticks.y = element_blank(),
            plot.title = element_text(hjust = 0.5),
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p7 <- ggplot(df3, aes(x = date, y = avg_funding_rate * 100)) +
      geom_col(aes(fill = positive), alpha = 0.7) +
      geom_hline(yintercept = 0, linetype = "solid") +
      geom_hline(yintercept = c(0.01, 0.05, -0.01),
                 linetype = "dashed", color = "gray50", alpha = 0.5) +
      scale_fill_manual(values = c("TRUE" = "green", "FALSE" = "red")) +
      labs(title = "Funding Rates",
           x = "Date", y = "Funding Rate (%)") +
      theme_minimal() +
      theme(legend.position = "none",
            plot.title = element_text(hjust = 0.5),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9))

p8 <- ggplot(df4, aes(x = date)) +
      geom_line(aes(y = oi_percentile, color = "OI Percentile"), linewidth = 0.5) +
      geom_line(aes(y = funding_percentile, color = "Funding Percentile"),
                linewidth = 0.5, linetype = "dashed") +
      geom_hline(yintercept = c(10, 90), linetype = "dotted", color = "gray50") +
      scale_color_manual(values = c("OI Percentile" = "blue",
                                    "Funding Percentile" = "red")) +
      labs(title = "Percentile Ranks",
           x = "Date", y = "Percentile (%)",
           color = "Metric") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            legend.position = "right",
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9))

p9 <- ggplot(m2_data, aes(x = date, y = m2_yoy)) +
      geom_line(color = "darkgreen", linewidth = 0.5) +
      geom_hline(yintercept = c(0, 5), linetype = "dashed", color = "red") +
      annotate("rect", xmin = min(m2_data$date), xmax = max(m2_data$date),
               ymin = 5, ymax = max(m2_data$m2_yoy, na.rm = TRUE),
               alpha = 0.1, fill = "green") +
      annotate("rect", xmin = min(m2_data$date), xmax = max(m2_data$date),
               ymin = min(m2_data$m2_yoy, na.rm = TRUE), ymax = 0,
               alpha = 0.1, fill = "red") +
      labs(title = "M2 Money Supply Growth",
           x = "Date", y = "YoY % Change") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

p10 <- ggplot(dxy_df, aes(x = date, y = dxy)) +
      geom_line(color = "blue", linewidth = 0.5) +
      # FIXED: Thresholds for FRED's DTWEXBGS index
      geom_hline(yintercept = c(110, 120), linetype = "dashed", color = "red") +
      annotate("rect", xmin = min(dxy_df$date), xmax = max(dxy_df$date),
               ymin = 110, ymax = 120, alpha = 0.1, fill = "orange") +
      labs(title = "US Dollar Index (DTWEXBGS)",
           x = "Date", y = "Index Value") +
      theme_minimal() +
      theme(plot.title = element_text(hjust = 0.5),
            legend.text = element_text(size = 7),       
            legend.title = element_text(size = 7),       
            legend.key.size = unit(0.25, "cm"),           
            legend.spacing.x = unit(0.2, "cm"),
            axis.title.x = element_text(size = 9),
            axis.title.y = element_text(size = 9)) + 
      scale_x_date(
        date_breaks = "6 months",
        date_labels = "%b %Y"
      )

# First create the left and right stacks
left_stack <- p1 / p2 / p3 / p9 / p10
right_stack <- p4 / p5 / p6 / p7 / p8

# Then combine them side by side and apply layout
combined_plot <- (left_stack | right_stack)

# Display the plot
combined_plot

```
